---
title: "Eli's Bets"
author: "Derek Nedveck"
date: "7/23/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

```


## Intro

During a car ride back from the North Shore of Minnesota, my friend Eli told my of his betting strategy:

> "I don't go into casinos often in my life, let's say at most 8 times. So each time that I go in, I'm going to go to the roulette table, place down a bet on a color, and if I win, I leave with my winnings. If I lose, then I double my bet on a single color, and go again. I keep doing this until I win, and then I leave."

My first instinct was "the house always wins", and to not try the game. But then my second instict was "I can simulate this in R to better understand this strategy."

So here is that simulation.

## Simulating the process

I'm going to write a function to simulate the process of Eli going to play one round of his betting strategy. One thing that I left out was that we did discuss one thing that would stop him would be havin a certain maximum amount to bet (and therefore maximum losses), so I'll need to capture that in the function. Eli is going to have a certain starting bet (like \$2), and have e.g. \$5K on hand to play the game with.

Here's a function to simulate that process:

```{r}
# writing a function for spinning a wheel
SpinWheel <- function(odds_win){
  sample(c(TRUE, FALSE), size = 1, prob = c(odds_win, 1 - odds_win))
}

PlayBettingGame <- function(starting_bet, cash_pool, odds_win){
  
  winnings <- 0
  bet <- starting_bet
  
  # going to do the initial round outside the while loop
  winnings <- winnings - bet
  cash_pool <- cash_pool - bet
  
  if(
    # this is a shortcut to not need to store a value
    SpinWheel(odds_win)
  ){
    winnings <- winnings + bet*2
    return(winnings)
  }

  # while loop to play this game until we win, or our cash pool is gone
  win <- FALSE
  while(win == FALSE){
    # place bets
    bet <- bet*2
    cash_pool <- cash_pool - bet
    if(cash_pool < 0){
      return(winnings)
    }
    winnings <- winnings - bet
    
    # spin wheel
    win <- SpinWheel(odds_win)
    # handle outcomes
    if(win == TRUE){
      winnings <- winnings + bet*2
    }
  }
  return(winnings)
}
```


```{r}
tracker <- 0 
w = F
while(w == F){
  tracker = tracker + 1
  w = rnorm(1) > 0
}
tracker
```

```{r}
run_track <- function(odds_win){
  tracker <- 0 
  w = F
  while(w == F){
    tracker = tracker + 1
    w = runif(1) < odds_win
  }
  return(tracker)
}
replicate(n = 20, expr = run_track(0.75))
```



```{r}
replicate(n = 20, expr = PlayBettingGame(2, 4000, 0.5))
```

