---
title: "Evaluating Eli's strategy"
author: "Derek"
date: "2022-07-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library(tibble)
library(ggplot2)
library(dplyr)
```


## Intro

As my friend Eli and I were driving back from the North Shore of Lake Superior, Eli filled me in on a betting strategy that he thought of:

> "I don't go into a casino often in my life, maybe 8 times all together, so each time I go into one, I'm going to go to a roulette table, bet on a single color (let's say black), and if I lose, double my bet and play again. I'm going to do this until I win, and then leave once I win. There's a chance I could win right away, and I would leave then."

My first thought was "the house always wins, this strategy won't work," but then I was thinking that I could simulate this in R to better understand how his strategy would perform. 

This is that simulation.

## Modeling the process

I created the following function to simulate the strategy that Eli came up with:

```{r}
# first, I need to spin a wheel

SpinWheel <- function(odds_win){
  runif(1) < odds_win
}

# function to simulate the double-until-win strategy
# odds_win = the odds of a win
# start_bet = the amount that one starts with
# cash_pool = maximum amount that one will bet with

RunStrategy <- function(odds_win, start_bet, cash_pool){
  # see how many attempts before a win
  win <- FALSE
  wtracker <- 0
  while(win == FALSE){
    wtracker <- wtracker + 1
    win <- SpinWheel(0.5)
  }
  wtracker
  
  # determine the amount of money spent on bets
  bet_money <- sum(start_bet*2^((wtracker-1):0))
  
  if(cash_pool > 0 ){
    if(bet_money > cash_pool){
      return(
        # if the attempts exhausts the cash pool, let's just pretend we 
        # bet the most we could
        - sum(start_bet*2^(
          (floor(log(cash_pool / start_bet, 2))-1):0)
          )
      )
    }
  }
  
  # determining the bet payout
  payout <- start_bet*2^wtracker
  
  # find total winnings
  winnings <- payout - bet_money
  
  return(winnings)
}
```


*Note: This is actually a boring strategy in the sense that a single run has only two outcomes -- winning the starting bet, or losing the max they can based on their available money pool.*

## Running simulations 

### A single run

Let's say Eli strolls into the casino with $5K to try out his strategy, what's the probability that Eli will leave with a positive value of winnings? 

First, what's the prob of winning in American Roulette?

Eli's going to bet on one color (red / black), which has the probability of winning being 18/ 38 (36 alternating red/black numbers, plus a 0 and 00 square). Since we're doing simulations, let's be exact and use the fraction.

I'm going to use the `replicate()` function to run my function call many times, and then I can start asking questions of those outcomes. 

```{r}
set.seed(333)
sr_reps <- 90003
single_run <- replicate(sr_reps, RunStrategy(odds_win = 18/38, start_bet = 2, cash_pool = 5000))
```

Alright, so what is the Pr[winnings > 0]?

```{r}
sum(single_run > 0) / sr_reps
```

And what's the Pr[winnings < 0]? 

```{r}
sum(single_run < 0) / sr_reps
```


### Eli's life plan

Alright, as previously described, Eli is thinking that he will only be going into a casino 8 times in his life. So now the thing to replicate is running his strategy 8 times, and then seeing how the breakdown is for what he could win during those simulated lifetimes. 

```{r}
lf_reps <- 90003
life_run <- replicate(lf_reps, {
  sum(
    replicate(n = 8, 
              RunStrategy(odds_win = 18/38, start_bet = 2, cash_pool = 5000)
                )
  )
  
})
```

Alright, what are the outcomes that Eli had?

```{r}
lf_table <- table(life_run)
lf_table
```

Wow, in Eli's life, he has the chance to win $16 ... and the crowd goes mild. 

Let's view this in probabilities

```{r}
lf_tbl <- tibble(
  winnings = names(lf_table),
  probability = as.vector(lf_table) / lf_reps
)
lf_tbl
```


So, I guess not too bad, 99.6% of winning $16 over 8 runs. I guess this means that you can beat the house.



### Cranking the bets up

I think one thing to explore is how the probability of winning shifts as your bets go up, and keeping a realistic cash pool.

Let's go back to simulating a single session at the roulette table, keep the cash pool at $5k, but then vary the bets, and see how the prob of winning changes.

```{r}
br_reps <- 90003
bets <- c(10, 20, 50, 100, 250, 500, 1000, 2000)
bet_ls <- lapply(bets, function(bet){
  replicate(br_reps, RunStrategy(odds_win = 18/38, start_bet = bet, cash_pool = 5000))
})
```

Now lets find the probability of winning for each 

```{r}
bet_probs <- bet_ls %>% lapply(function(run){
  sum(run > 0) / br_reps
  }) %>% 
  unlist

bet_tbl <- tibble(
  starting_bet = bets,
  prob_positive_winning = bet_probs
)
bet_tbl
```

So now it comes down to how much risk you are willing to tolerate to get a 100% gain on your bet. Maybe \$250 or \$500? Or maybe do something else with the $5K that one would need to play with. 



